---
title: "Assignment5"
author: "Neelima Cherukuri"
date: '2022-06-07'
output: html_document
---

```{r}
library(XML)
library(xml2)
library(RCurl)
library(dplyr)
library(stringr)
library(tidyverse)
library(ggplot2)
library(tidyr)
library(lubridate)
library(scales)
library(readr)
```

Question 1 :
Load the XML data directly from the URL below into a data frame (or tibble) and display the dimensions
of the data.
URL: https://www.senate.gov/general/contact_information/senators_cfm.xml.
If you get errors while loading the XML file, ensure that you use the RCurl package. Refer to this post for
details.
```{r}
fileurl <- "https://www.senate.gov/general/contact_information/senators_cfm.xml"
download.file(fileurl, destfile = "senators_cfm.xml")
Senators <- xmlToDataFrame("senators_cfm.xml")
dim(Senators)
str(Senators)



```

#Explanation: This dataframe has 101 rows and 13 columns

Question 2:
Construct a regular expression (regex) to extract only the first, last names, and party (D,R,I) of each 
senator; the regex should remove their middle initial and/or suffix e.g. remove Jr. III, etc. Ensure that the 
updated names are reflected in the dataframe. 

```{r}
#Extracting names and removing middle name and initials

new_first_name <- gsub('([A-z]+).*', '\\1', Senators$first_name)
new_first_name

new_last_name <- gsub('([A-z]+).*', '\\1', Senators$last_name)
new_last_name

#Replacing the updated column values

Senators[["first_name"]] <- new_first_name
colnames(Senators)[colnames(Senators)== "first_name"] <- "first_name"

Senators[["last_name"]] <- new_last_name
colnames(Senators)[colnames(Senators)== "last_name"] <- "last_name"

#The data frame has been updated
str(Senators)

```

Question 3:
Create a function called senatorsByState() which takes the two letter abbreviation for a US State as its 
input argument and displays the first name, last name and party of each senator for the selected state. 
For example, if the user enters ‘MA’ your function should display a message that is similar to the 
following: “The senators for MA are: Edward Markey, Democratic Party and Elizabeth Warren, 
Democratic Party”
```{r}
senatorsByState <- function(x) 
  {
  abc <- Senators %>%
    select(state, first_name, last_name, party) %>%
    group_by(state) %>%
    filter(state == (x))
    output <- paste("The senators for",(x), "are: ")
  name_change <- paste(abc$first_name, abc$last_name, ",", recode(abc$party,"D" = "Democratic Party", "R" = "Republician Party", "I" = "Independent Party"))
  add <- str_c(name_change, collapse = " and ")
  combined <- str_c(output, add)
  return(combined)
}

#Calling the new function
senatorsByState("IL")
senatorsByState("CO")


```

Question 4:
Download the attached csv file from Canvas and load it in your R environment. Perform steps to tidy the 
data and the prepared data should be divided across two tibbles named country_name and 
indicator_data. The country_name tibble should contain the country name and country code (ensure 
that you remove duplicates), and the indicator_data tibble should include the country_code, year, and 
value. Note: Tidy the data using pivot_longer(), pivot_wider() and separate(), where applicable. 
```{r}
a <- read.csv("C:\\Users\\neelc\\Downloads\\Ratio Of Female To Male Youth Unemployment Rate .csv", header = TRUE, skip=4, sep = ",")
view(a)

#Tidying the data
tidy <- pivot_longer(a, c(X1960:X2020), names_to = "year", values_to = "value", values_drop_na  = TRUE) %>%
separate(year, into = c("x","year"), sep = "X") %>%
select(-Indicator.Name,-Indicator.Code,-X,-x)
#Tibble one created from original dataset a

country_tibble <- a %>%
   select (Country.Name, Country.Code)
country_tibble
#Detecting duplicate values. 0 duplicate values
country_tibble[!duplicated(country_tibble), ]
view(country_tibble)

#Tibble two created from the tidy dataset
indicator_data <- tidy %>% 
  select("Country.Code", "year", "value")
view(indicator_data)

```
Question 5:
Select five countries from each of the following continents: North America, Asia and Middle East. 
Visualize their respective data from the last 20 years using a line chart; use facet_wrap to display the data 
for each continent in its own chart. Explain your observations about the ratio of female to male youth 
unemployment rate in the selected regions. 
```{r}

b <- read.csv("C:\\Users\\neelc\\Downloads\\Country Meta-Data.csv")
View(b)

#Creating my own tibble and inner joining with the previously made one 
mytable <-  tibble(Country.Name= c("Ethiopia","Guinea", "Iraq", "Liberia", "Lesotho", "China", "India", "Bhutan", "Indonesia", "Sri Lanka", "Albania", "Germany", "Italy", "Spain", "Denamrk"),continent= c("Africa", "Africa", "Africa", "Africa", "Africa", "Asia", "Asia", "Asia","Asia", "Asia", "Europe", "Europe", "Europe", "Europe", "Europe")) %>%
  inner_join(country_tibble, by= "Country.Name") %>%
  inner_join(indicator_data, by= "Country.Code") %>%
  filter(year == (2001:2020))

View(mytable)

#Plotting line grapg
ggplot(data= mytable, aes(x= year, y= value)) +
  geom_line() +
  facet_wrap(.~continent,) +
  theme(axis.text.x = element_text(angle= 90, hjust=0)) +
  ggtitle("F:M unemployment comparison between continents") +
  xlab("Years") + ylab("F:M Ratio")

```

Comment- The graph is a mix 3 different continents. It shows that For Asia the ratio remains more or less unchanged during the years. For Africa the ratio is low and has been deteriorating over the past 6-7 years. Europe also has a similar case, The unemployment ratio has been up and down and decreased over the past few years.