---
title: "Assignment6"
author: "Neelima Cherukuri"
date: "6/27/2022"
output: html_document
---

```{r}
library(RSQLite)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(plyr)
```


```{r}
db <- dbConnect(SQLite(), dbname= "/Users/harshavardhanreddymogulla/Downloads/imdb.db")
dbListTables(db)
```

#Loading imdb.db in R console
sqlite> .open "C:/Users/neelc/Downloads/imdb.db" 

sqlite> .tables

#1 table found 

movie_info 

sqlite> select * from movie_info;

sqlite> .mode column

sqlite> .headers on

sqlite> select * from movie_info;

#Q1 Create a table named director_info using SQLite; the columns are: Director_ID, and Director_Name. The Director_ID should be the primary key. 

sqlite DROP TABLE director_info;

sqlite> CREATE TABLE director_info (Director_ID INT PRIMARY KEY, Director_Name TEXT NOT NULL);

sqlite> .tables

#new table created
director_info  movie_info  

# Q2- Import the entire data from the CSV file into the director_info table using the SQLite .importcommand (see helpful resources below). Verify that your data was imported correctly. 
sqlite .separator ','
sqlite> .mode csv director_info
sqlite> .import /Users/maitrishah/Downloads/directors.csv --skip 1 director_info
sqlite> select * from director_info ORDER BY Director_ID;

```{r}
#Q2
#1. Count the number of rows in the movie_info and director_info tables. 
dbGetQuery(db, "SELECT count(*) from director_info")  #Returns 548, therefore number of rows in table director_info is equal to 548
dbGetQuery(db, "SELECT count(*) from movie_info")  #Returns 1000, therefore number of rows in table movie_info is equal
```

```{r}
#Q2
#2. How many movies were released between 2010 and 2020 (inclusive)? Visualize the results. 
release <- dbGetQuery(db, "SELECT * FROM movie_info
           WHERE Release_Year >= 2010 and Release_Year <= 2020")
#248 movies were released between 2010 and 2020 (inclusive)

# Visualizing the results.
ggplot(data= release, aes(x= Release_Year, y= ..count.., fill= Release_Year)) +
  geom_bar()+
  geom_text(stat= "count", aes(label = ..count..))+
   labs(title= "Movies released between 2010 and 2020 (inclusive)",
        caption = "Chart to visulaize movie count",
        fill= "Year of Release") +
   xlab("Year") + ylab("No. of movies")
  
#Year 2014 has the most released movies (32) followed by 2013 and 2016 (28 each) while the lease number of movies were released in the year 2020 (6)
```


```{r}
#Q2
#3 What is the minimum, average and maximum ratings for “Action” movies.  Ensure that you query the genre using wild cards. 
action <- dbGetQuery(db, "SELECT * from movie_info WHERE Genre LIKE '%Action%'") #Using wilcard LIKE
rating <- dbGetQuery(db, "SELECT min(IMDB_Rating) as Min_rating, avg(IMDB_Rating) as Avg_rating, max(IMDB_Rating) as Max_rating from movie_info WHERE Genre LIKE '%Action%'")

#The minimum rating for Action movies is 7.6 , maximum is 9 with an average rating of ~7.95
```


```{r}
#Q2.
#4 What are the 25 highest-grossing movies within the dataset? Display the title, genre and gross.

highest <- dbGetQuery(db, "Select Series_Title, Genre, Gross from movie_info WHERE Gross != 'NA' ORDER BY Gross") #Getting list of movies after removing NA values in the gross column and arranging it in ascending order

tail(highest, 25) #Getting the 25 highest grossing movies
#Star Wars: Episode VII - The Force Awakens is the top-most grossing movie of Genre Action, Adventure, Sci-Fi and earning a gross amount of $936662225. The lowest among the top 25 is Joker of Genre Crime, Drama, Thriller with an earning of $335451311
```




```{r}
#Q2.5 Which directors have the highest-grossing movies. Display the director name and the total gross. Ensure that you join the necessary tables. Visualize the results using a Bar chart. 
 #Joining the director_info and movie_info
Join_Table <- dbGetQuery(db, "Select di.Director_Name, mo.Gross from movie_info mo
                 join director_info di
                 on di.Director_ID= mo.Director_ID
                 where Gross != 'NA'
                 order by Gross")
head(Join_Table)

#using library dplyr to calculate total gross for each director
Calc_Gross <- ddply(Join_Table,"Director_Name",numcolwise(sum)) %>% 
  arrange(desc(Gross))

colnames(Calc_Gross)[2] <- "Total_Gross"  #Changing column name 2 to Total Gross
Top_Gross <- head(Calc_Gross, 25) #Top 25 directors
Top_Gross

# Visualizing the results using a Bar chart
ggplot(data= Top_Gross, aes(x= reorder(Director_Name, -Total_Gross), y= Total_Gross)) +
  geom_histogram(stat= "identity") +
  theme(axis.text.x = element_text(angle= 90, hjust=0))+
   labs(title= "Top 25 directors",
        subtitle= "Directors with highest total gross",
        caption = "Chart to visulaize the top 25 directors") +
   xlab("Director names") + ylab("Total Gross($)")

#The director with the highest gross is Steven Spielberg with a total gross of $2478133165 while the lowest among the top 25 is Gareth Edwards with a total gross of $532177324
```



```{r}
#Q2.6  Create a function called verifyDirector() that takes a director name as its argument, and queries the database to check if the director exists. Your function should display a message to notify the user if the director was found or not. 

verifyDirector <- function(x)
{
input <- dbSendQuery(db, "SELECT * FROM director_info WHERE Director_Name = :x")
dbBind(input, params = list(x = x))
output <- dbFetch(input)
if (nrow(output) < 1)
{
print(paste("Oops! Director not found, please check again"))
}
else 
{
dbClearResult(input, output)   
combined <- str_c(output$Director_Name, paste(" -Bingo, the director you entered is present in the dataset"))
return(combined)
dbClearResult(combined)
}}

#To verify the created function
verifyDirector('Aamir Khan')
verifyDirector('Neelima Cherukuri')
```



