---
title: "Assignment9"
author: "Neelima Cherukuri"
date: '2022-07-23'
output: html_document
---

```{r}
library(tidyverse)
library(dplyr)
```

Question 1:
Loading the data
```{r}
# Download the data from the URL
url <- "https://data.sfgov.org/api/views/rkru-6vcg/rows.csv?accessType=DOWNLOAD"

download_file <- download.file(url,
              destfile ="\\Users\\neelc\\Downloads\\Air_Traffic_Passenger_Statistics.csv" ,
              method = "curl")

# Load the file
data <- read.csv("C:\\Users\\neelc\\Downloads\\Air_Traffic_Passenger_Statistics.csv", header=TRUE)
str(data)
```

Question 2:

```{r}
# Add the column year and month after extracting them from the Activity.Period
data$year <- substring(data$Activity.Period,1,4)

data$month <- substring(data$Activity.Period,5,6)

# Filter the month of March and extract all domestic passenger activity for each year
graph_data <- data%>%
  filter(month == "03" &  GEO.Summary == "Domestic")%>%
  group_by(year, Activity.Period)%>%
  summarize(Total_Passengers = sum(Passenger.Count))
graph_data
ggplot(graph_data, aes(x=year, y=Total_Passengers, group = 1 )) + 
  geom_line() + labs(title = "Line chart for the domestic passenger activity that occured each year in March")
```

Question 3:

```{r}
# Convert the year column from a character vector to numeric
graph_data$year <- as.numeric(graph_data$year)

new_data <- graph_data %>%
  select(year, Total_Passengers)%>%
  filter_all(all_vars(year %in% c(2016,2017,2018)))

# Moving Average
# Calculate the 3-Year Moving Average for the following time periods : 201603, 201703 and 201803
# tail(new_data)

moving_avg_forecast <- mean(new_data$Total_Passengers)
moving_avg_forecast
# Create a dataframe for the forecasted value 
moving_avg_forecast_new <- data.frame(moving_avg_forecast)
# class(moving_avg_forecast_new)

# Calculate the error by subtracting the actual and the predicted values for March 2019
6899726	 - 6771607
```
#Comment:The predicted value for total passenger activity for March 2019 is 6771607 which is close enough to the actual value of 6899726 The error after subtracting the actual and predicted value is 128119. Moreover,the predicted value is slightly lower than the actual value.
Question 4:

```{r}
# Use three year Weighted Moving Average to forecast the total passenger acticity for 2019 for the following time periods : 201603, 201703 and 201803
# Apply the weights 3,5,7

weights <- c(3,5,7)

sw <- weights*(new_data$Total_Passengers)
sw
Final_calculation <- sum(sw)/sum(weights)
Final_calculation
# Create a dataframe for the forecasted value 
weighted_moving_avg_forecast <- data.frame(Final_calculation)
# class(weighted_moving_avg_forecast)

# Calculate the error by subtracting the actual and the predicted values for March 2019
6899726 - 6865349
```
#Comment:The predicted value for total passenger activity for March 2019 is 6899726 which is close to the actual value of 6865349.The error after subtracting the actual and predicted value is 34377. Moreover,the predicted value is slightly lower than the actual value.

Question 5:


```{r}
# Exponential Smoothing Model
# Select the data from the years 2008 to 2018
new_data_1 <- graph_data %>%
  select(year, Total_Passengers)%>%
  filter_all(all_vars(year %in% c(2008,2009,2010,2011,2012,2013,2014,2015,2016,2017,2018)))
# Define alpha 
a <- 0.7

# Define the forecast column and the error column
new_data_1$FT <- 0
new_data_1$E <- 0

# Calculate the first forecast value manually
new_data_1$FT[1]<-new_data_1$Total_Passengers[1]
# head(new_data_1)
# new_data_1$FT[2]<-new_data_1$FT[1]+a*new_data_1$E[1]

# calculate the error - (observed value - forecasted value)
new_data_1$E[1]<-new_data_1$Total_Passengers[1]-new_data_1$FT[1]

# Create a loop to calculate all the forecast values
for(i in 2:nrow(new_data_1)){
  new_data_1$FT[i]<-new_data_1$FT[i-1] + a * new_data_1$E[i-1]
  new_data_1$E[i]<-new_data_1$Total_Passengers[i]-new_data_1$FT[i]
}
new_data_1
# Calculate the prediction for March 2019
n<-nrow(new_data_1)
F.estimate<-new_data_1$FT[n]+a*new_data_1$E[n]
F.estimate
# Create a dataframe for the forecasted value 
exponential_smoothing_model_forecast <- data.frame(F.estimate)
# class(exponential_smoothing_model_forecast)
```

Question 6:

```{r}
# Select the data from the years 2008 to 2018
new_data_2 <- graph_data %>%
  select(year, Total_Passengers)%>%
  filter_all(all_vars(year %in% c(2008,2009,2010,2011,2012,2013,2014,2015,2016,2017,2018)))

# Build a simple linear regression model using the year and total passenger activity for all data from 2008 to 2018. 
# Select the data from the years 2008 to 2018
model <- lm(new_data_2$Total_Passengers ~ new_data_2$year)
summary(model)
print(model)
# Forecast the total passenger activity for 2019 and 2020.
F_2019 <-  --524789117  + 263538 * (2019)
F_2019
F_2020 <-  --524789117  + 263538 * (2020)
F_2020
# Create a dataframe for the forecasted value 
linear_regression_model_forecast <- data.frame(F_2019)
# class(linear_regression_model_forecast)
```



Question 7:

```{r}
# MSE for the exponential smoothing model (Question 5)

# Define alpha 
a <- 0.7

# Define the forecast column and the error column
new_data_1$FT <- 0
new_data_1$E <- 0

# Calculate the first forecast value manually
new_data_1$FT[1]<-new_data_1$Total_Passengers[1]
# head(new_data_1)
# new_data_1$FT[2]<-new_data_1$FT[1]+a*new_data_1$E[1]

# calculate the error - (observed value - forecasted value)
new_data_1$E[2]<-new_data_1$Total_Passengers[2]-new_data_1$FT[2]

# Create a loop to calculate all the forecast values
for(i in 2:nrow(new_data_1)){
 new_data_1$FT[i]<-new_data_1$FT[i-1] + a * new_data_1$E[i-1]
  new_data_1$E[i]<-new_data_1$Total_Passengers[i]-new_data_1$FT[i]
}
new_data_1
# Calculate the MSE for exponential smoothing model
MSE_exponential <- mean(new_data_1$E)
MSE_exponential
#  MSE for the linear regression  model (Question 6)

# Build the model based on the data from 2008 to 2018
model <- lm(new_data_2$Total_Passengers ~ new_data_2$year)
# summary(model)

print(model)
# Define the forecast column and the error column
new_data_2$FT_value <- 0
new_data_2$absErr <- 0

# head(new_data_2)

# Calculate the first forecasted value
# new_data_2$FT_value[1]<- 131769*new_data_2$Total_Passengers[1] -262394559  
#head(new_data_2)

# Calculate the first error
# new_data_2$absErr[1]<-abs(new_data_2$Total_Passengers - new_data_2$FT_value[1])
# head(new_data_2)

# Create a loop to calculate all the forecast values and the Error
for(i in 1:nrow(new_data_2)){
  new_data_2$FT_value[i]<-263538*new_data_2$Total_Passengers[i]-524789117       
  new_data_2$absErr[i]<-abs(new_data_2$Total_Passengers - new_data_2$FT_value[i])
}
new_data_2
#  Calculate the MSE for the linear regression model
MSE_linear <- mean(new_data_2$absErr)
MSE_linear
```
#Comment:THE MSE for exponential smoothing model is 299280.9 THE MSE for linear regression model is  1.504953e+12 The exponential smoothing model has a smaller MSE.
